Parsing Postfix log files

### 2009-02-12 doc/logparser-acronyms.tex
Fixed acronyms so they work properly, get expanded when appropriate, and add the     correct entries to the glossary and acronym appendices. Correct capitalisation of most section and subsection headings. Corrections made to background chapter; more remain to be made.

\newglossaryentry{PLP}{
    type=\acronymtype,
    name={PLP},
    description={Postfix Log Parser\glsadd{PLP glossary}}
}
\newglossaryentry{PLP glossary}{
    name={Postfix Log Parser},
    description={
        Postfix Log Parser is the parser implementing the algorithm
        described in this thesis.
    }
}

\newglossaryentry{DNSBL}{
    type=\acronymtype,
    name={DNSBL},
    description={DNS Blacklist\glsadd{DNSBL glossary}}
}
\newglossaryentry{DNSBL glossary}{
    name={DNS Blacklist},
    description={
        A DNS Blacklist~\cite{Wikipedia-DNSBL} is a simple collaborative
        anti-spam technique used to reject or penalise email sent from mail
        servers reported to have sent large volumes of spam.  To use a
        DNSBL the mail server makes a DNS request incorporating the IP
        address of the client; if the requested hostname exists the client
        is on the DNSBL, and the mail server can decide what course of
        action to take.
    }
}

\newglossaryentry{API}{
    type=\acronymtype,
    name={API},
    description={Application Programming Interface\glsadd{API glossary}}
}
\newglossaryentry{API glossary}{
    name={Application Programming Interface},
    description={
        One of the fundamental concepts when writing programs is the reuse
        of existing code, so that each program does not reinvent the wheel.
        When a body of code is intended to be reused repeatedly, the user
        of this code needs to be informed of the functionality provided by
        the code.  An API defines the interface provided to the user, and
        acts as a contract between the user and the provider: if the user
        adheres to the API the provider guarantees it will work, but is
        free to change the implementation provided the API is preserved.
    }
}

\newglossaryentry{SMTP}{
    type=\acronymtype,
    name={SMTP},
    description={Simple Mail Transfer Protocol\glsadd{SMTP glossary}}
}
\newglossaryentry{SMTP glossary}{
    name={Simple Mail Transfer Protocol},
    description={
        SMTP is the protocol that transfers mail from the sender to the
        recipient across the Internet.  It is a simple, human readable,
        plain text protocol, making it quite simple to test and debug
        problems with it.  A detailed description of SMTP is beyond the
        scope of this document: the original protocol definition is
        RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821};
        introductory guides can be found
        at~\cite{smtp-intro-01,smtp-intro-02}.
    }
}

\newglossaryentry{LMTP}{
    type=\acronymtype,
    name={LMTP},
    description={Local Mail Transfer Protocol\glsadd{LMTP glossary}}
}
\newglossaryentry{LMTP glossary}{
    name={Local Mail Transfer Protocol},
    description={
        LMTP is a protocol derived from SMTP that removes the need for the
        server to maintain a mail delivery queue, instead relying on the
        client to maintain it.  Typically the client is an MTA, and the
        server a delivery agent or a mail store.  Full details are
        available in~\cite{lmtp-rfc}.
    }
}

\newglossaryentry{MTA}{
    type=\acronymtype,
    name={MTA},
    description={Mail Transfer Agent\glsadd{MTA glossary}}
}
\newglossaryentry{MTA glossary}{
    name={Mail Transfer Agent},
    description={
        A Mail Transfer Agent sends and receives mail via SMTP\@.  Users
        submit mail to an MTA via their mail client (e.g.\ Microsoft
        Outlook, Thunderbird, webmail services); the sending MTA transfers
        the mail to the receiving MTA (which may in turn forward it to
        another MTA).
    }
}

\newglossaryentry{RFC}{
    type=\acronymtype,
    name={RFC},
    description={Request For Comments\glsadd{RFC glossary}}
}
\newglossaryentry{RFC glossary}{
    name={Request For Comments},
    description={
        The Request For Comments series is a series of proposals defining
        various protocols and file formats, e.g.\ SMTP\@.  The name is
        somewhat misleading: initially the authors were asking for peer
        review, but these documents are now the de facto standards the
        Internet runs on.
    }
}

\newglossaryentry{SQL}{
    type=\acronymtype,
    name={SQL},
    description={Structured Query Language\glsadd{SQL glossary}}
}
\newglossaryentry{SQL glossary}{
    name={Structured Query Language},
    description={
        The standard language used for database querying, modification, and
        maintenance.  Some information and history can be found at
        \url{http://en.wikipedia.org/wiki/SQL}; a good introduction can be
        found at \url{http://philip.greenspun.com/sql/}.
    }
}

\newglossaryentry{LMA}{
    type=\acronymtype,
    name={LMA},
    description={Log Mail Analyzer\glsadd{LMA glossary}}
}
\newglossaryentry{LMA glossary}{
    name={Log Mail Analyzer},
    description={
        One of the other Postfix log file parsers reviewed and the only
        example of prior published paper~\cite{log-mail-analyser} found
        that specifically deals with parsing Postfix log files.
    }
}

\newglossaryentry{CSV}{
    type=\acronymtype,
    name={CSV},
    description={Comma-Separated Value\glsadd{CSV glossary}}
}
\newglossaryentry{CSV glossary}{
    name={Comma-Separated Value},
    description={
        The most basic form of database available, each record is a single
        line in the file and the fields are separated by a special
        character, typically a comma or colon.
    }
}

\newglossaryentry{FQDN}{
    type=\acronymtype,
    name={FQDN},
    description={Fully Qualified Domain Name\glsadd{FQDN glossary}}
}
\newglossaryentry{FQDN glossary}{
    name={Fully Qualified Domain Name},
    description={
        An FQDN is a hostname plus domain name, e.g.\ \newline{}
            example.com     is a domain name          \newline{}
            www             is a hostname             \newline{}
            www.example.com is a FQDN
    }
}

\newglossaryentry{pid}{
    type=\acronymtype,
    name={pid},
    description={Process Identifier\glsadd{pid glossary}}
}
\newglossaryentry{pid glossary}{
    name={Process Identifier},
    description={
        There may be multiple copies of any program executing at any one
        time, so the program's name is not suitable as a distinguishing
        identifier; instead each process executing is given a pid that is
        guaranteed to be unique for the lifetime of the process.  Once the
        process has completed, the pid may be reused, as they are drawn
        from a finite pool.
    }
}

\newglossaryentry{UCE}{
    type=\acronymtype,
    name={UCE},
    description={Unsolicited Commercial Email\glsadd{UCE glossary}}
}
\newglossaryentry{UCE glossary}{
    name={Unsolicited Commercial Email},
    description={
        UCE is a more restrictive definition of spam than most people use:
        it only covers mail that is explicitly commercial, thus excluding
        viruses, Bayesian poisoning mails, backscatter, and those annoying
        chain letters you get from friends.
    }
}

\newglossaryentry{ESMTP}{
    type=\acronymtype,
    name={ESMTP},
    description={Extended SMTP\glsadd{ESMTP glossary}}
}
\newglossaryentry{ESMTP glossary}{
    name={Extended SMTP},
    description={
        ESMTP is Extended SMTP, defined in RFC~1869~\cite{RFC1869}.
    }
}

\newglossaryentry{IP}{
    type=\acronymtype,
    name={IP},
    description={Internet Protocol\glsadd{IP glossary}}
}
\newglossaryentry{IP glossary}{
    name={Internet Protocol},
    description={
        The Internet Protocol is the protocol used to communicate between
        computers on the Internet.  An IP address is a unique address
        assigned to a computer on the Internet, allowing it to communicate
        with other computers on the Internet.
    }
}

\newglossaryentry{SLCT}{
    type=\acronymtype,
    name={SLCT},
    description={Simple Logfile Clustering Tool\glsadd{SLCT glossary}}
}
\newglossaryentry{SLCT glossary}{
    name={Simple Logfile Clustering Tool},
    description={
        SLCT~\cite{slct-paper} is a tool implementing an algorithm designed
        by Risto Vaarandi for identifying, grouping, and producing a regex
        to match similar log lines.
    }
}

\newglossaryentry{ATN}{
    type=\acronymtype,
    name={ATN},
    description={Augmented Transition Networks\glsadd{ATN glossary}}
}
\newglossaryentry{ATN glossary}{
    name={Augmented Transition Networks},
    description={
        Originally described in~\cite{atns} and further in~\cite{nlpip},
        Augmented Transition Networks are a tool used in Computational
        Linguistics for creating grammars to parse or generate sentences,
    }
}

\newglossaryentry{CLI}{
    type=\acronymtype,
    name={CLI},
    description={Command Line Interface\glsadd{CLI glossary}}
}
\newglossaryentry{CLI glossary}{
    name={Command Line Interface},
    description={
        An computer interface based on typing commands, rather than using a
        mouse.
    }
}

\newglossaryentry{NLP}{
    type=\acronymtype,
    name={NLP},
    description={Natural Language Processing\glsadd{NLP glossary}}
}
\newglossaryentry{NLP glossary}{
    name={Natural Language Processing},
    description={
        Natural Language Processing~\cite{Wikipedia-nlp} attempts to
        increase our understanding of the languages normally used by humans
        (e.g.\ English, Spanish, Japanese), with the goal of writing
        programs that can understand human languages.
    }
}

>>>>

\newglossaryentry{PLP}{
    type=\acronymtype,
    name={PLP},
    description={Postfix Log Parser},
    first={Postfix Log Parser (PLP)}
}
\newglossaryentry{PLP glossary}{
    name={Postfix Log Parser},
    description={
        Postfix Log Parser is the parser implementing the algorithm
        described in this thesis.
    }
}

\newglossaryentry{DNSBL}{
    type=\acronymtype,
    name={DNSBL},
    description={DNS Blacklist},
    first={DNS Blacklist (DNSBL)}
}
\newglossaryentry{DNSBL glossary}{
    name={DNS Blacklist},
    description={
        A DNS Blacklist~\cite{Wikipedia-DNSBL} is a simple collaborative
        anti-spam technique used to reject or penalise email sent from mail
        servers reported to have sent large volumes of spam.  To use a
        DNSBL the mail server makes a DNS request incorporating the IP
        address of the client; if the requested hostname exists the client
        is on the DNSBL, and the mail server can decide what course of
        action to take.
    }
}

\newglossaryentry{API}{
    type=\acronymtype,
    name={API},
    description={Application Programming Interface},
    first={Application Programming Interface (API)}
}
\newglossaryentry{API glossary}{
    name={Application Programming Interface},
    description={
        One of the fundamental concepts when writing programs is the reuse
        of existing code, so that each program does not reinvent the wheel.
        When a body of code is intended to be reused repeatedly, the user
        of this code needs to be informed of the functionality provided by
        the code.  An API defines the interface provided to the user, and
        acts as a contract between the user and the provider: if the user
        adheres to the API the provider guarantees it will work, but is
        free to change the implementation provided the API is preserved.
    }
}

\newglossaryentry{SMTP}{
    type=\acronymtype,
    name={SMTP},
    description={Simple Mail Transfer Protocol},
    first={Simple Mail Transfer Protocol (SMTP)}
}
\newglossaryentry{SMTP glossary}{
    name={Simple Mail Transfer Protocol},
    description={
        SMTP is the protocol that transfers mail from the sender to the
        recipient across the Internet.  It is a simple, human readable,
        plain text protocol, making it quite simple to test and debug
        problems with it.  A detailed description of SMTP is beyond the
        scope of this document: the original protocol definition is
        RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821};
        introductory guides can be found
        at~\cite{smtp-intro-01,smtp-intro-02}.
    }
}

\newglossaryentry{LMTP}{
    type=\acronymtype,
    name={LMTP},
    description={Local Mail Transfer Protocol},
    first={Local Mail Transfer Protocol (LMTP)}
}
\newglossaryentry{LMTP glossary}{
    name={Local Mail Transfer Protocol},
    description={
        LMTP is a protocol derived from SMTP that removes the need for the
        server to maintain a mail delivery queue, instead relying on the
        client to maintain it.  Typically the client is an MTA, and the
        server a delivery agent or a mail store.  Full details are
        available in~\cite{lmtp-rfc}.
    }
}

\newglossaryentry{MTA}{
    type=\acronymtype,
    name={MTA},
    description={Mail Transfer Agent},
    first={Mail Transfer Agent (MTA)}
}
\newglossaryentry{MTA glossary}{
    name={Mail Transfer Agent},
    description={
        A Mail Transfer Agent sends and receives mail via SMTP\@.  Users
        submit mail to an MTA via their mail client (e.g.\ Microsoft
        Outlook, Thunderbird, webmail services); the sending MTA transfers
        the mail to the receiving MTA (which may in turn forward it to
        another MTA).
    }
}

\newglossaryentry{RFC}{
    type=\acronymtype,
    name={RFC},
    description={Request For Comments},
    first={Request For Comments (RFC)}
}
\newglossaryentry{RFC glossary}{
    name={Request For Comments},
    description={
        The Request For Comments series is a series of proposals defining
        various protocols and file formats, e.g.\ SMTP\@.  The name is
        somewhat misleading: initially the authors were asking for peer
        review, but these documents are now the de facto standards the
        Internet runs on.
    }
}

\newglossaryentry{SQL}{
    type=\acronymtype,
    name={SQL},
    description={Structured Query Language},
    first={Structured Query Language (SQL)}
}
\newglossaryentry{SQL glossary}{
    name={Structured Query Language},
    description={
        The standard language used for database querying, modification, and
        maintenance.  Some information and history can be found at
        \url{http://en.wikipedia.org/wiki/SQL}; a good introduction can be
        found at \url{http://philip.greenspun.com/sql/}.
    }
}

\newglossaryentry{LMA}{
    type=\acronymtype,
    name={LMA},
    description={Log Mail Analyzer},
    first={Log Mail Analyzer (LMA)}
}
\newglossaryentry{LMA glossary}{
    name={Log Mail Analyzer},
    description={
        One of the other Postfix log file parsers reviewed and the only
        example of prior published paper~\cite{log-mail-analyser} found
        that specifically deals with parsing Postfix log files.
    }
}

\newglossaryentry{CSV}{
    type=\acronymtype,
    name={CSV},
    description={Comma-Separated Value},
    first={Comma-Separated Value (CSV)}
}
\newglossaryentry{CSV glossary}{
    name={Comma-Separated Value},
    description={
        The most basic form of database available, each record is a single
        line in the file and the fields are separated by a special
        character, typically a comma or colon.
    }
}

\newglossaryentry{FQDN}{
    type=\acronymtype,
    name={FQDN},
    description={Fully Qualified Domain Name},
    first={Fully Qualified Domain Name (FQDN)}
}
\newglossaryentry{FQDN glossary}{
    name={Fully Qualified Domain Name},
    description={
        An FQDN is a hostname plus domain name, e.g.\ \newline{}
            example.com     is a domain name          \newline{}
            www             is a hostname             \newline{}
            www.example.com is a FQDN
    }
}

\newglossaryentry{pid}{
    type=\acronymtype,
    name={pid},
    description={Process Identifier},
    first={Process Identifier (pid)}
}
\newglossaryentry{pid glossary}{
    name={Process Identifier},
    description={
        There may be multiple copies of any program executing at any one
        time, so the program's name is not suitable as a distinguishing
        identifier; instead each process executing is given a pid that is
        guaranteed to be unique for the lifetime of the process.  Once the
        process has completed, the pid may be reused, as they are drawn
        from a finite pool.
    }
}

\newglossaryentry{UCE}{
    type=\acronymtype,
    name={UCE},
    description={Unsolicited Commercial Email},
    first={Unsolicited Commercial Email (UCE)}
}
\newglossaryentry{UCE glossary}{
    name={Unsolicited Commercial Email},
    description={
        UCE is a more restrictive definition of spam than most people use:
        it only covers mail that is explicitly commercial, thus excluding
        viruses, Bayesian poisoning mails, backscatter, and those annoying
        chain letters you get from friends.
    }
}

\newglossaryentry{ESMTP}{
    type=\acronymtype,
    name={ESMTP},
    description={Extended SMTP},
    first={Extended SMTP (ESMTP)}
}
\newglossaryentry{ESMTP glossary}{
    name={Extended SMTP},
    description={
        ESMTP is Extended SMTP, defined in RFC~1869~\cite{RFC1869}.
    }
}

\newglossaryentry{IP}{
    type=\acronymtype,
    name={IP},
    description={Internet Protocol},
    first={Internet Protocol (IP)}
}
\newglossaryentry{IP glossary}{
    name={Internet Protocol},
    description={
        The Internet Protocol is the protocol used to communicate between
        computers on the Internet.  An IP address is a unique address
        assigned to a computer on the Internet, allowing it to communicate
        with other computers on the Internet.
    }
}

\newglossaryentry{SLCT}{
    type=\acronymtype,
    name={SLCT},
    description={Simple Logfile Clustering Tool},
    first={Simple Logfile Clustering Tool (SLCT)}
}
\newglossaryentry{SLCT glossary}{
    name={Simple Logfile Clustering Tool},
    description={
        SLCT~\cite{slct-paper} is a tool implementing an algorithm designed
        by Risto Vaarandi for identifying, grouping, and producing a regex
        to match similar log lines.
    }
}

\newglossaryentry{ATN}{
    type=\acronymtype,
    name={ATN},
    description={Augmented Transition Networks},
    first={Augmented Transition Networks (ATN)}
}
\newglossaryentry{ATN glossary}{
    name={Augmented Transition Networks},
    description={
        Originally described in~\cite{atns} and further in~\cite{nlpip},
        Augmented Transition Networks are a tool used in Computational
        Linguistics for creating grammars to parse or generate sentences,
    }
}

\newglossaryentry{CLI}{
    type=\acronymtype,
    name={CLI},
    description={Command Line Interface},
    first={Command Line Interface (CLI)}
}
\newglossaryentry{CLI glossary}{
    name={Command Line Interface},
    description={
        An computer interface based on typing commands, rather than using a
        mouse.
    }
}

\newglossaryentry{NLP}{
    type=\acronymtype,
    name={NLP},
    description={Natural Language Processing},
    first={Natural Language Processing (NLP)}
}
\newglossaryentry{NLP glossary}{
    name={Natural Language Processing},
    description={
        Natural Language Processing~\cite{Wikipedia-nlp} attempts to
        increase our understanding of the languages normally used by humans
        (e.g.\ English, Spanish, Japanese), with the goal of writing
        programs that can understand human languages.
    }
}

### 2009-02-12 doc/logparser-architecture.tex
Fixed acronyms so they work properly, get expanded when appropriate, and add the     correct entries to the glossary and acronym appendices. Correct capitalisation of most section and subsection headings. Corrections made to background chapter; more remain to be made.

\section{Parser architecture and design}

@@

There is some similarity between the parser's design and William Wood's
\gls{ATN}~\cite{atns,nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
\gls{ATN} and the parser is accidental, but it is obvious that the two
different approaches share an underlying separation of concerns; this
appears to be a natural division of responsibility and functionality.
\Tableref{Similarities between ATN and this architecture} shows the
similarities.

@@

\begin{table}[ht]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tabletopline{}%
        \gls{ATN}          & Architecture  & Similarity                \\
        \tablemiddleline{}%
        Networks        & Algorithm     & Determines the sequence 
                                        of transitions              \\
                        &               & or actions that 
                                        constitutes a valid  input. \\
        Transitions     & Actions       & Save data and impose
                                        conditions the              \\
                        &               & input must meet to be
                                        accepted as valid.          \\
        Abbreviations   & Rules         & Responsible for 
                                        classifying input.          \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

\subsection{Rule characteristics}

@@

\subsection{Overlapping rules}

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the order in which rules
are tried against each line may change between log files, and the first
matching rule wins.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating \gls{SMTP} delivery to
specific sites from \gls{SMTP} delivery to the rest of the world.  The
algorithm provides a mechanism for ordering overlapping rules: the priority
field in each rule (defaults to zero).  Negative priorities may be useful
for catchall rules.

@@

    \item Sort by regex and visually inspect the list, e.g.\ with \gls{SQL}
        similar to: \textbf{select regex from rules order by regex;}

>>>>

\section{Parser Architecture and Design}

@@

There is some similarity between the parser's design and William Wood's
\acronym{ATN}~\cite{atns,nlpip}, a tool used in Computational Linguistics
for creating grammars to parse or generate sentences.  The resemblance
between \acronym{ATN} and the parser is accidental, but it is obvious that
the two different approaches share an underlying separation of concerns;
this appears to be a natural division of responsibility and functionality.
\Tableref{Similarities between ATN and this architecture} shows the
similarities.

@@

\begin{table}[ht]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tabletopline{}%
        \acronym{ATN}   & Architecture  & Similarity                \\
        \tablemiddleline{}%
        Networks        & Algorithm     & Determines the sequence 
                                        of transitions              \\
                        &               & or actions that 
                                        constitutes a valid  input. \\
        Transitions     & Actions       & Save data and impose
                                        conditions the              \\
                        &               & input must meet to be
                                        accepted as valid.          \\
        Abbreviations   & Rules         & Responsible for 
                                        classifying input.          \\
        \tablebottomline{}%
    \end{tabular}
\end{table}

@@

\subsection{Rule Characteristics}

@@

\subsection{Overlapping Rules}

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the order in which rules
are tried against each line may change between log files, and the first
matching rule wins.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating \acronym{SMTP} delivery to
specific sites from \acronym{SMTP} delivery to the rest of the world.  The
algorithm provides a mechanism for ordering overlapping rules: the priority
field in each rule (defaults to zero).  Negative priorities may be useful
for catchall rules.

@@

    \item Sort by regex and visually inspect the list, e.g.\ with
        \acronym{SQL} similar to: \textbf{select regex from rules order by
        regex;}
